Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-02 article-models-tt.tex
introducing lemmas about composition; big code cleanup

\begin{align*}
& \inference{\jtype{\Gamma}{A}}{\jfam{\Gamma}{A}} & & \inference{\jtypeeq{\Gamma}{A}{B}}{\jfameq{\Gamma}{A}{B}}\\
& \inference{\jtype{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jtype{\Gamma}{B}}
& & \inference{\jtype{\Gamma}{A}\quad\jfameq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{A}{B}}
\end{align*}

\subsection{The basic rules for judgmental equality}
The rules for judgmental equality establish that it is an equivalence relation
in all three cases (contexts, types and terms).
\bgroup\small
\begin{align*}
& \inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} 
& & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} 
& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\greek{E}}}{\jctxeq{\Gamma}{\greek{E}}}\\
& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{A}{A}} 
& & \inference{\jfameq{\Gamma}{A}{B}}{\jfameq{\Gamma}{B}{A}}
& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\Gamma}{B}{C}}{\jfameq{\Gamma}{A}{C}}\\
& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}}
& & \inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}}
& & \inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
\end{align*}
\egroup

@@

\begin{align*}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfam{\Gamma}{A}}{\jfam{\Delta}{A}}
& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jfameq{\Delta}{A}{B}}\\
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
& \inference{\jfameq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
\end{align*}

@@

\begin{align}
& \inference{}{\jctx{\emptyc}}\\
& \inference{\jctx{\Gamma}}{\jfam{\Gamma}{\emptyf[\Gamma]}}\\
& \inference{\jctx{\Gamma}}{\jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
& \inference{\jterm{\Gamma}{\emptyf[\Gamma]}{x}}{\jtermeq{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
\end{align}

@@

\begin{align}
& \inference{\jctx{\Gamma}}{\jfam{\emptyc}{\Gamma}} 
& & \inference{\jfam{\emptyc}{\Gamma}}{\jctx{\Gamma}}\\
& \inference{\jctxeq{\Gamma}{\Delta}}{\jfameq{\emptyc}{\Gamma}{\Delta}}
& & \inference{\jfameq{\emptyc}{\Gamma}{\Delta}}{\jctxeq{\Gamma}{\Delta}}
\end{align}

\subsubsection{The empty context is compatible with itslef}
The empty context $\emptyc$ may be considered as a family of contexts over the empty
context. When we do this, we get $\emptyf[\emptyc]$.
\begin{equation}
\inference{}{\jfameq{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
\end{equation}
In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
guarantees that this will not cause confusion. Likewise, we shall denote
$\emptytm[\Gamma]$ by $\emptytm$.

@@

\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
& \inference{\jfam{{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{align}

\subsubsection{Extension is compatible with the empty context}
The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
\begin{align}
& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
& \inference{\jctx{\Gamma}}{\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
\end{align}

@@

\begin{align}
& \inference{\jfam{\Gamma}{A}\qquad\jfam{{\Gamma}{A}}{P}}
  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
& \inference{\jfam{{\Gamma}{A}}{P}\qquad\jfam{{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
\end{align}

@@

 In the following inference rules we assume that $\jfam{\Gamma}{A}$ and in the
 rules asserting a judgmental equality we assume furthermore that 
 $\jfameq{\Gamma}{A}{A'}$.
\begin{align}
& \inference{\jfam{\Gamma}{B}}{\jfam{{\Gamma}{A}}{\ctxwk{A}{B}}}
& & \inference{\jfameq{\Gamma}{B}{B'}}{\jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
& \inference{\jfam{{\Gamma}{B}}{Q}}
{\jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
& & \inference{\jfameq{{\Gamma}{B}}{Q}{Q'}}
{\jfameq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
& \inference{\jterm{\Gamma}{B}{y}}{\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
& & \inference{\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
& \inference{\jterm{{\Gamma}{B}}{Q}{g}}{\jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
& & \inference{\jtermeq{{\Gamma}{B}}{Q}{g}{g'}}
{\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}{\ctxwk{A'}{g'}}}
\end{align}

@@

\begin{align}
& \inference{\jfam{\Gamma}{B}\quad\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
\end{align}

\subsubsection{Weakening is compatible with the empty context}
The following rules express that when the empty context or context family is
weakened, the result is the empty context family.
\begin{align}
& \inference{\jctx{\Gamma}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
& \inference{\jfam{\Gamma}{A}}{\jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
\end{align}
Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
their terms, families over those families and
terms of those unchanged:
\begin{align}
& \inference{\jfam{\Gamma}{B}}{\jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}}\\
& \inference{\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
& \inference{\jfam{{\Gamma}{B}}{Q}}{\jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
& \inference{\jterm{{\Gamma}{B}}{Q}{g}}{\jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
\end{align}

@@

\subsubsection{Weakening is compatible with itself}\label{comp-ww}
We state judgmental equality rules expressing
that weakening is compatible with itself. These rules state that the following
diagram commutes given any two families $A$ and $B$ in context $\Gamma$:
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\jfam{\Gamma}{\blank} \ar{r}{C\mapsto\ctxwk{B}{C}} \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} & \jfam{{\Gamma}{B}}{\blank} \ar{d}{Q\mapsto\ctxwk{A}{Q}}\\
\jfam{{\Gamma}{A}}{\blank} \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} & \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
\end{tikzcd}
\end{equation*}
Thus, we get the following set of inference rules:
\begin{align}
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{{\Gamma}{C}}{R}}
          {\jfameq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
            {\ctxwk{A}{{B}{R}}}
            {\ctxwk{{A}{B}}{{A}{R}}}}\label{comp-ww-f}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{{\Gamma}{C}}{R}{t}}
          {\jtermeq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
            {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
            \label{comp-ww-t}
\end{align}

@@

In the following inference rules we assume that
$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. 
\begin{align}
& \inference{\jfam{{\Gamma}{B}}{Q}}
    {\jfameq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\label{comp-ew-f}\\
& \inference{\jterm{{\Gamma}{B}}{Q}{g}}
    {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\label{comp-ew-t}
\end{align}

@@

\begin{align}
& \inference{\jfam{{\Gamma}{A}}{P}}{\jfam{\Gamma}{\subst{x}{P}}}
& & \inference{\jfameq{{\Gamma}{A}}{P}{P'}}{\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
& \inference{\jfam{{{\Gamma}{A}}{P}}{Q}}{\jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
& & \inference{\jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}}{\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
& & \inference{\jtermeq{{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
& \inference{\jterm{{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
& &\inference{\jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
\end{align}

We also add the following rule asserting that substitution preserves the
property of being a type:
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}\quad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
\end{align}

\subsubsection{Substitution is compatible with the empty context}
The fibers of the empty family are the empty families:
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
\end{align}

The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
leaves everything unchanged.
\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\subst{\emptytm}{A}}{A}}\\
& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}}\\
& \inference{\jfam{{\Gamma}{A}}{P}}{\jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
& \inference{\jterm{{\Gamma}{A}}{P}{f}}{\jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
\end{align}

\subsubsection{Substitution is compatible with extension}\label{comp-se}
Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.
\begin{align}
& \inference{\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
  {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
\end{align}

>>>>

\begin{align*}
& \inference
  {\jtype{\Gamma}{A}}
  {\jfam{\Gamma}{A}}
& & \inference
    {\jtypeeq{\Gamma}{A}{B}}
    {\jfameq{\Gamma}{A}{B}}\\
& \inference
  {\jtype{\Gamma}{A}
   \jfameq{\Gamma}{A}{B}}
  {\jtype{\Gamma}{B}}
& & \inference
    {\jtype{\Gamma}{A}
     \jfameq{\Gamma}{A}{B}}
    {\jtypeeq{\Gamma}{A}{B}}
\end{align*}

\subsection{The basic rules for judgmental equality}
The rules for judgmental equality establish that it is an equivalence relation
in all three cases (contexts, types and terms).
\bgroup\small
\begin{align*}
& \inference
  {\jctx{\Gamma}}
  {\jctxeq{\Gamma}{\Gamma}} 
& & \inference
    {\jctxeq{\Gamma}{\Delta}}
    {\jctxeq{\Delta}{\Gamma}} 
& & \inference
    {\jctxeq{\Gamma}{\Delta}
     \jctxeq{\Delta}{\greek{E}}}
    {\jctxeq{\Gamma}{\greek{E}}}\\
& \inference
  {\jfam{\Gamma}{A}}
  {\jfameq{\Gamma}{A}{A}} 
& & \inference
    {\jfameq{\Gamma}{A}{B}}
    {\jfameq{\Gamma}{B}{A}}
& & \inference
    {\jfameq{\Gamma}{A}{B}
     \jfameq{\Gamma}{B}{C}}
    {\jfameq{\Gamma}{A}{C}}\\
& \inference
  {\jterm{\Gamma}{A}{x}}
  {\jtermeq{\Gamma}{A}{x}{x}}
& & \inference
    {\jtermeq{\Gamma}{A}{x}{y}}
    {\jtermeq{\Gamma}{A}{y}{x}}
& & \inference
    {\jtermeq{\Gamma}{A}{x}{y}
     \jtermeq{\Gamma}{A}{y}{z}}
    {\jtermeq{\Gamma}{A}{x}{z}}
\end{align*}
\egroup

@@

\begin{align*}
& \inference
  {\jctxeq{\Gamma}{\Delta}
   \jfam{\Gamma}{A}}
  {\jfam{\Delta}{A}}
& & \inference
    {\jctxeq{\Gamma}{\Delta}
     \jfameq{\Gamma}{A}{B}}
    {\jfameq{\Delta}{A}{B}}\\
& \inference
  {\jctxeq{\Gamma}{\Delta}
   \jterm{\Gamma}{A}{x}}
  {\jterm{\Delta}{A}{x}}
& & \inference
    {\jctxeq{\Gamma}{\Delta}
     \jtermeq{\Gamma}{A}{x}{y}}
    {\jtermeq{\Delta}{A}{x}{y}}\\
& \inference
  {\jfameq{\Gamma}{A}{B}
   \jterm{\Gamma}{A}{x}}
  {\jterm{\Gamma}{B}{x}}
& & \inference
    {\jfameq{\Gamma}{A}{B}
     \jtermeq{\Gamma}{A}{x}{y}}
    {\jtermeq{\Gamma}{B}{x}{y}}
\end{align*}

@@

\begin{align}
& \inference
  {}
  {\jctx{\emptyc}}\\
& \inference
  {\jctx{\Gamma}}
  {\jfam{\Gamma}{\emptyf[\Gamma]}}\\
& \inference
  {\jctx{\Gamma}}
  {\jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}}\\
& \inference
  {\jterm{\Gamma}{\emptyf[\Gamma]}{x}}
  {\jtermeq{\Gamma}{\emptyf[\Gamma]}{x}{\emptytm[\Gamma]}}
\end{align}

@@

\begin{align}
& \inference
  {\jctx{\Gamma}}
  {\jfam{\emptyc}{\Gamma}} 
& & \inference
    {\jfam{\emptyc}{\Gamma}}
    {\jctx{\Gamma}}\\
& \inference
  {\jctxeq{\Gamma}{\Delta}}
  {\jfameq{\emptyc}{\Gamma}{\Delta}}
& & \inference
    {\jfameq{\emptyc}{\Gamma}{\Delta}}
    {\jctxeq{\Gamma}{\Delta}}
\end{align}

\subsubsection{The empty context is compatible with itslef}
The empty context $\emptyc$ may be considered as a family of contexts over the empty
context. When we do this, we get $\emptyf[\emptyc]$.
\begin{equation}
\inference
{}
{\jfameq{\emptyc}{\emptyc}{\emptyf[\emptyc]}}
\end{equation}
In the future, we shall denote $\emptyf[\Gamma]$ by $\emptyf$. The above rule
guarantees that this will not cause confusion. Likewise, we shall denote
$\emptytm[\Gamma]$ by $\emptytm$.

@@

\begin{align}
& \inference
  {\jfam{\Gamma}{A}}
  {\jctx{\ctxext{\Gamma}{A}}}
& & \inference
    {\jctxeq{\Gamma}{\Delta}
     \jfameq{\Gamma}{A}{B}}
    {\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfam{\Gamma}{\ctxext{A}{P}}}
& & \inference
    {\jfameq{\Gamma}{A}{B} 
     \jfameq{{\Gamma}{A}}{P}{Q}}
    {\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{align}

\subsubsection{Extension is compatible with the empty context}
The following rule asserts that extension by $\emptyc$ leaves the contexts unchanged.
\begin{align}
& \inference
  {\jctx{\Gamma}}
  {\jctxeq{\ctxext{\emptyc}{\Gamma}}{\Gamma}}\\
& \inference
  {\jctx{\Gamma}}
  {\jctxeq{\ctxext{\Gamma}{\emptyf}}{\Gamma}}\\
& \inference
  {\jfam{\Gamma}{A}}
  {\jfameq{\Gamma}{\ctxext{\emptyf}{A}}{A}}
\end{align}

@@

\begin{align}
& \inference
  {\jfam{\Gamma}{A}
   \jfam{{\Gamma}{A}}{P}}
  {\jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}
   \jfam{{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{A}{{P}{Q}}}}
\end{align}

@@

 In the following inference rules we assume that $\jfam{\Gamma}{A}$ and in the
 rules asserting a judgmental equality we assume furthermore that 
 $\jfameq{\Gamma}{A}{A'}$.
\begin{align}
& \inference
  {\jfam{\Gamma}{B}}
  {\jfam{{\Gamma}{A}}{\ctxwk{A}{B}}}
& & \inference
    {\jfameq{\Gamma}{B}{B'}}
    {\jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
& \inference
  {\jfam{{\Gamma}{B}}{Q}}
  {\jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
& & \inference
    {\jfameq{{\Gamma}{B}}{Q}{Q'}}
    {\jfameq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
& \inference
  {\jterm{\Gamma}{B}{y}}
  {\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
& & \inference
    {\jtermeq{\Gamma}{B}{y}{y'}}
    {\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}\\
& \inference
  {\jterm{{\Gamma}{B}}{Q}{g}}
  {\jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}}
& & \inference
    {\jtermeq{{\Gamma}{B}}{Q}{g}{g'}}
    {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}{\ctxwk{A'}{g'}}}
\end{align}

@@

\begin{align}
& \inference
  {\jfam{\Gamma}{B}
   \jtype{\ctxext{\Gamma}{B}}{Q}}
  {\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
\end{align}

\subsubsection{Weakening is compatible with the empty context}
The following rules express that when the empty context or context family is
weakened, the result is the empty context family.
\begin{align}
& \inference
  {\jctx{\Gamma}}
  {\jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}}\\
& \inference
  {\jfam{\Gamma}{A}}
  {\jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}}
\end{align}
Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
their terms, families over those families and
terms of those unchanged:
\begin{align}
& \inference
  {\jfam{\Gamma}{B}}
  {\jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}}\\
& \inference
  {\jterm{\Gamma}{B}{y}}
  {\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
& \inference
  {\jfam{{\Gamma}{B}}{Q}}
  {\jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
& \inference
  {\jterm{{\Gamma}{B}}{Q}{g}}
  {\jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
\end{align}

@@

\subsubsection{Weakening is compatible with itself}\label{comp-ww}
We state judgmental equality rules expressing
that weakening is compatible with itself. These rules state that the following
diagram commutes given any two families $A$ and $B$ in context $\Gamma$:
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\jfam{\Gamma}{\blank} \ar{r}{C\mapsto\ctxwk{B}{C}} \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} & \jfam{{\Gamma}{B}}{\blank} \ar{d}{Q\mapsto\ctxwk{A}{Q}}\\
\jfam{{\Gamma}{A}}{\blank} \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} & \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
\end{tikzcd}
\end{equation*}
Thus, we get the following set of inference rules:
\begin{align}
& \inference
  {\jfam{\Gamma}{A}
   \jfam{\Gamma}{B}
   \jfam{{\Gamma}{C}}{R}}
  {\jfameq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
    {\ctxwk{A}{{B}{R}}}
    {\ctxwk{{A}{B}}{{A}{R}}}}\label{comp-ww-f}\\
& \inference
  {\jfam{\Gamma}{A}
   \jfam{\Gamma}{B}
   \jterm{{\Gamma}{C}}{R}{t}}
  {\jtermeq{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
    {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
\label{comp-ww-t}
\end{align}

@@

In the following inference rules we assume that
$\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. 
\begin{align}
& \inference
  {\jfam{{\Gamma}{B}}{Q}}
  {\jfameq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}
  \label{comp-ew-f}\\
& \inference
  {\jterm{{\Gamma}{B}}{Q}{g}}
  {\jtermeq{{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}} 
  \label{comp-ew-t}
\end{align}

@@

\begin{align}
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfam{\Gamma}{\subst{x}{P}}}
& & \inference
    {\jfameq{{\Gamma}{A}}{P}{P'}}
    {\jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
& \inference
  {\jfam{{{\Gamma}{A}}{P}}{Q}}
  {\jfam{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
& & \inference
    {\jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}}
    {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
& \inference
  {\jterm{{\Gamma}{A}}{P}{f}}
  {\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
& & \inference
    {\jtermeq{{\Gamma}{A}}{P}{f}{f'}}
    {\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
& \inference
  {\jterm{{{\Gamma}{A}}{P}}{Q}{g}}
  {\jterm{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
& & \inference
    {\jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}}
    {\jtermeq{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
\end{align}

We also add the following rule asserting that substitution preserves the
property of being a type:
\begin{align}
& \inference
  {\jterm{\Gamma}{A}{x}
   \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}
  {\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
\end{align}

\subsubsection{Substitution is compatible with the empty context}
The fibers of the empty family are the empty families:
\begin{align}
& \inference
  {\jterm{\Gamma}{A}{x}}
  {\jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jfam{{\Gamma}{A}}{P}}
  {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\emptyf}}{\emptyf}}
\end{align}

The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
leaves everything unchanged.
\begin{align}
& \inference
  {\jfam{\Gamma}{A}}
  {\jfameq{\Gamma}{\subst{\emptytm}{A}}{A}}\\
& \inference
  {\jterm{\Gamma}{A}{x}}
  {\jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}}\\
& \inference
  {\jterm{{\Gamma}{A}}{P}{f}}
  {\jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}}.
\end{align}

\subsubsection{Substitution is compatible with extension}\label{comp-se}
Suppose $\jterm{\Gamma}{A}{x}$ in all of the following inference rule.
\begin{align}
& \inference
  {\jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
  {\jfameq{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxext{Q}{R}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}}
\end{align}

\subsubsection{Substitution is compatible with weakening}\label{comp-sw}
The rules asserting the compatibility of substitution with weakening assert
that the following diagram commutes for any $\jterm{\Gamma}{A}{x}$ and any
$\jfam{{\Gamma}{A}}{P}$:
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\jfam{{\Gamma}{A}}{\blank} \ar{d}[swap]{Q\mapsto\subst{x}{Q}} \ar{r}{Q\mapsto\ctxwk{P}{Q}} & \jfam{{{\Gamma}{A}}{P}}{\blank} \ar{d}{R\mapsto\subst{x}{R}}\\ 
\jfam{\Gamma}{\blank} \ar{r}[swap]{B\mapsto\ctxwk{\subst{x}{P}}{B}} & \jfam{{\Gamma}{\subst{x}{P}}}{\blank}
\end{tikzcd}
\end{equation*}
We plug in an extra layer of families to cover the most general case at once.
In the following inference rules we assume that $\jterm{\Gamma}{A}{x}$ and
$\jfam{{\Gamma}{A}}{P}$:
\begin{align}
& \inference
  {\jfam{{{\Gamma}{A}}{Q}}{R}}
  {\jfameq{{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}{\subst{x}{\ctxwk{P}{R}}}{\ctxwk{\subst{x}{P}}{\subst{x}{R}}}}
  \label{comp-sw-f}\\
& \inference
  {\jterm{{{\Gamma}{A}}{Q}}{R}{h}}
  {\jtermeq
      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
      {\subst{x}{\ctxwk{P}{R}}}
      {\subst{x}{\ctxwk{P}{h}}}
      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}}
  \label{comp-sw-t}
\end{align}
